Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-09-01 main.tex
corrected review typos

\begin{example}\label{ex:SWT}
We build a small \SWT over the alphabets $\Sigma$ and $\Deltai$
(Ex.~\ref{ex:running} and \ref{distance-time}),
%weighted transducer model
with two states $q_0$ and $q_1$, that
calculates the temporal distance between an input performance in~$\Sigma^*$
and the subsequence
of $\Deltai$ events in a score.
%the performance $I$ and
%the score $O$ of Example~\ref{ex:running}.
%
Given  a performed event $\mu$
and the corresponding notated event $\nu$
(\eg MIDI pitch 69 and note $\mbox{\footnotesize A4}$),
the weight  computed by the \SWT is the time distance between both,
as modeled by transitions $\wei_{11}$ below.
%
The continuation symbol $'-'$
(met for instance in \emph{ties}
$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
or \emph{dots} \musQuarterDotted{})
%that represents continuation
is  skipped with no cost (transitions $\wei_{01}$).
%Likewise, the markups are ignored.
\[
\begin{array}{lclclcll}
\wei_{11}(q_0, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) & \quad & %|\tau' - \tau| & \quad &
\wei_{11}(q_1, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) &
     \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
\wei_{01}(q_0, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one\\
\wei_{10}(q_0, \mu\ts{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also want to take performing errors into account,
since a performer could, for example, play an unwritten extra note.
%\lydia{reformulated this sentence}
%
%Reading such a note is modelled by
The transition $\wei_{10}$,
with an fixed weight value $\alpha \in \Semiring$,
switches from state $q_0$ (normal) to $q_1$ (error)
when reading an extra note $\mu$.
%\philippe{Comprends pas cette phrase}
%\florent{$\to$ reformulation}
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and
$\init(q_1) = \final(q_1) = \zero$.
%\florent{dernière phrase utile?}
%This \SWT is capable of evaluating the differences between a score and a performance,
%all the while ensuring that performance errors are plausible.
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

@@

\noindent
 \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined  as the transducers of Definition~\ref{def:SWT},
by simply omitting the output symbols.
%
%In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$,
, respectively $\mathsf{out} : Q \to \Semiring$,  %\Phi_\Sigma,$
are functions defining the weight for entering
(respectively leaving) computation in a state,
and ${\wei_1}$ is a transition function
from $Q \times Q$ into~$\Phi_{\Sigma}$.
\end{definition}
%
\noindent
%\florent{mv formal def. weight for \SWA to app.?}
As above in the case of \SWT,
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$,
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$.
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$,
The computation of $A$ on words $s \in \Sigma^*$
is based on with an intermediate function $\weight_A$,
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$
%
\begin{align}
\weight_A(q, \varepsilon, q') & = \one\ \mathrm{if}\  q = q'\ 
          \mathrm{and}\  \zero\  \mathrm{otherwise}\\
%\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q}
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$
\begin{equation}
A(s)  =
\displaystyle\bigoplus_{q, q' \in Q} \init(q)
\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q')
\label{eq:SWA-value}
\end{equation}

@@

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}\label{sec:SWVPA}
The model presented now generalizes symbolic \VPA
(\SVPA~\cite{dAntonyAlur14SVPDA},
 generalizing themselves \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
from Boolean semirings to arbitrary semiring domains.
It associates to every nested word over an infinite alphabet a weight value
in a semiring.
Nested words can describe structures of labeled trees.
In the context of parsing, they will be useful to
represent AST % parse trees.
(see Section~\ref{sec:parsing} and Appendix~\ref{app:trees}). % parse trees. 
%\florent{see \S 5 and App.A}

@@

\noindent
%The intuition is the following for the above transitions.
$\weiei$, $\weiec$, and $\weier$ describe the cases where the stack is empty.
%\lydia{moved this to the beginning}
%
\noindent
$\weii$ and $\weiei$ both read an input internal symbol $a$ and change state from $q$ to $q'$,
without changing the stack.
Moreover, $\weii$ reads a pair made of
${\call{c}} \in \Deltac$ and $p \in P$ on the top of the stack
($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$).
%
\noindent
$\weic$ and $\weiec$ read the input call symbol $\call{c}'$,
push it to the stack along with $p'$, and change state from $q$ to to $q'$.
Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack
($c$ is compared to $c'$).
%and $\weiec$ applies iff the stack is empty.
%
\noindent
$\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
Moreover, $\weir$ reads and pop from stack a pair made of $\call{c}$ and $p$,
($\call{c}$ is compared to $\return{r}$).
%and $\weier$ applies iff the stack is empty.
%In this case, the weight function $\phi_\mathsf{r}$
%computes a value of matching between the call and return symbols $c$ and $r$.
%This value might be set to $\zero$ in order to express that the symbols do not match.

Formally, the transitions of the automaton~$A$ are defined
with an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
%
A configuration $q[\gamma]$
is composed of a state $q \in Q$
and a stack content $\gamma \in \Gamma^*$,
where $\Gamma = \Deltac \times P$.
Hence, $\weight_A$ is a function from
$[Q \times \Gamma^*] \times \Delta^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
The empty stack is denoted by $\bot$, and the topupmost
symbol is the last pushed content.
%
The recursive definition of $\weight_A$
enumerates each of the six possible cases:
reading $a \in \Deltai$,
%\lydia{intro to func}
or $\call{c} \in \Deltac$, or $\return{r} \in \Deltar$,
for each possible state of the stack (empty or not).
%to add to $u \in {\Delta}^*$.
%\lydia{introduced the 6 cases}
\florent{shorter notation $cp$ for $\< c, p>$ ?}

@@

\noindent
Every $\SWA$ $A = \< Q, \init, {\wei_1}, \final >$,
over $\Sigma$, $\Semiring$ and $\bar\Phi$
is a particular case of $\SWVPA$
$\< Q, \emptyset, \init, \bar\wei, \final >$
over $\Delta$, $\Semiring$ and $\bar\Phi$
with $\Deltai = \Sigma$ and $\Deltac = \Deltar = \emptyset$,
and computing with an always empty stack:
$\weiei = \wei_1$ and all the other functions
of~$\bar\wei$ are the constant~$\zero$.

@@

%Let us assume a third component in the symbols of $\Deltac$
%\philippe{Je trouve l'exemple un peu confus. Je ne sais pas
%quel est "l'associate time interval" d'un symbole. Je ne sais
%pas à quoi fait référence "the following call transition". Je
%n'ose pas trop toucher mais je crois qu'il faut revoir.}
%which is the duration of the associated time interval.
%It is used to check the timestamps in the symbols.
%
For instance, the following call transition starts in state
$q_{{i}/{c}}$, meaning that the current interval has number~$i < c$
amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the stack top).
It reads a new time-division symbol~$\ccall{d}$:
\(
\weic\bigl(q_{{i}/{c}},
           \ccall{c}\ts{[\tau, \tau+\ell]}, q,
           \ccall{d}\ts{\iota}, q_{{i+1}/{c}},
           q_{{0}/{d}} \bigr) = \alpha_d
\).
The time interval $\iota$ attached to the~$\ccall{d}$ read 
must have a starting time $\tau + \frac{i\ell}{c}$,
where~$\tau$ and~$\ell$ are respectively
the starting time and duration of the interval attached to
the previous time-division symbol read $\ccall{c}$ (found on the stack).
%plus the total duration of intervals read so far, $i \ell$, and
%the new current interval duration is~$\frac{\ell}{d}$.
%
And it must have a duration $\frac{\ell}{d}$.

@@

%\subsection{Properties}
\medskip\noindent
Similarly to \VPA~\cite{AlurMadhusudan09nested}
and \SVPA~\cite{dAntonyAlur14SVPDA},
the class of \SWVPA is closed under the binary operators of the underlying semiring.
%
\begin{proposition}\label{prop:SWVPA-product}
Let $A_1$ and $A_2$ be two \SWVPA
over the same $\Delta$, commutative $\Semiring$ and effective $\bar\Phi$.
There exists two effectively constructible $\SWVPA$
$A_1 \oplus A_2$ and $A_1 \otimes A_2$,
such that for every $s \in \Delta^*$,
$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and
$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
\end{proposition}
%
\begin{proof}
We do a classical product construction, see Appendix~\ref{sec:product} for details.
%similar to the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
\end{proof}
\noindent
We present now a procedure for searching, for a~$\SWVPA$ $A$,
a word %$w \in \Delta^*$
of minimal weight for~$A$.
%as stated in the following proposition. %\wrt~$\leq_\oplus$,
%It requires the following property of label theories.
%
\begin{proposition}\label{th:best-search}
For a \SWVPA $A$
over $\Delta$,
$\Semiring$ commutative, bounded, total and complete, %semiring
and $\bar\Phi$ effective, %and $k$-convex, % label theory,
one can construct in PTIME a word $t \in \Delta^*$
such that $A(t)$ is minimal \wrt the natural ordering $\leq_\oplus$ for $\Semiring$.
\end{proposition}
%\florent{total?}
%
Let $A = \< Q, P, \init, \bar\wei, \final >$.
%
We propose a Dijkstra algorithm computing,
for every $q, q' \in Q$,
the minimum, \wrt $\leq_\oplus$, of the function
$\beta_{q, q'} : t \mapsto \weight_A(\config{q}{\bot}, t, \config{q'}{\bot})$.
Let us denote by $b_\bot(q, q')$ this minimum.
By definition of $\leq_\oplus$, and since $\Semiring$ is total,
it holds that:
%
\begin{equation}\label{eq:bbot}
  b_\bot(q, q') = \bigoplus_{t\in \Delta^*}
  \textstyle
  \weight_A\bigl(\config{q}{\bot}, t, \config{q'}{\bot}\bigr).
\end{equation}
The infinite sum in~\eqref{eq:bbot} is well defined since $\Semiring$ is complete.
%
Following~\eqref{eq:SWVPA-value}, and the associativity, commutativity
and distributivity for~$\otimes$ and~$\oplus$, the minimum of~$A(t)$ is:
\begin{equation}\label{eq:min}
{\displaystyle\bigoplus_{t\in \Delta^*}} A(t)
=
{\displaystyle\bigoplus_{t\in \Delta^*}}
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes}
\beta_{q, q'}(t)
\mathop{\otimes} \mathsf{out}(q')
=
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes}
b_\bot(q, q')
\mathop{\otimes} \mathsf{out}(q')
\end{equation}

\noindent
In order to compute the above function $b_\bot : Q \times Q \to \Semiring$,
we shall consider an auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
%
Intuitively, $b_\top(q, p, q')$ is a function of $\Phic$,
mapping every $c \in \Deltac$ to
the minimum weight of a computation of~$A$
starting in state $q$ with a non-empty stack
$\gamma' = \<c, p> \, \gamma \in \Gamma^+$,
and ending in state $q'$ with the same stack $\gamma'$,
such that moreover,
the computation does not pop the pair $\<c, p>$ at the top of $\gamma'$
(\ie $\gamma'$ is left untouched during the computation).
However, the computation can read $\<c, p>$ at the top of $\gamma'$,
and can also push another pair $\< c', p'> \in \Gamma$ on $\gamma'$,
following the third case of
in the definition \eqref{eq:SWVPA-weight} of $\weight_A$ (call symbol).
The pair $\< c', p'>$ can be pop later, during the computation from $q$ to $q'$,
following the fifth case of \eqref{eq:SWVPA-weight} (return symbol).
%However, it cannot apply one of the two last cases (return symbol and empty stack)
%when the current stack is $\gamma$.
%pop symbols in $\gamma$.
%
% Note that having a stack reduced to such a symbol makes impossible the application of the
% two last cases in the definition of $\weight_A$ (return symbol and empty stack).
% However, it is possible to apply the two first cases
% (internal symbol or call symbol, with a push on the top of $\top$).
%
Formally, in order to define $b_\top$, we consider
a fresh stack symbol $\top \notin \Gamma$,   %which does not belong to $\Gamma$,
representing the above untouched stack, and let:
%
\begin{equation}\label{eq:btop}
  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Delta^*}
  \textstyle
  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
\quad
\mathrm{~for~all~} c \in \Deltac
\end{equation}
%
By definition of $\weight_A$ in \eqref{eq:SWVPA-weight},
using the symbol $\top$ for the part of the stack below $\<c, p>$
(\ie the substack $\gamma$ in the above $\gamma' = \<c, p> \, \gamma$)
ensures that this part is not touched during the computation.
% there cannot be a pop of $\<c, p>$ followed by a push of $\<c, p>$
% because it is not possible to push over $\top$ (push read the top of stack).
%
This ensures in particular that the subword
read during the computation is well parenthesized
(every symbol in $\Deltac$ has a matching symbol in $\Deltar$).

@@

Let us now apply the models and results of the previous sections %in order to define
to the problem of parsing over an infinite alphabet. %appropriate
%
%\subsection{Definition}
%
Let~$\Sigma$
and~$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$
be countable input and output alphabets,
let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a
commutative, bounded, total and complete %\florent{total?} 
semiring and let $\bar\Phi$ be an effective label theory over $\Semiring$,
containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Deltai}$, as well as
$\Phii$, $\Phic$, $\Phir$, $\Phicr$
(following the notations of Section~\ref{sec:SWVPA-def}).
%
\noindent
We assume given the following input:
\begin{description}
\item[--] a \SWT $T$ over $\Sigma$, $\Deltai$, $\Semiring$, and $\bar\Phi$,
defining a measure %between words
$T: \Sigma^* \times \Deltai^* \to \Semiring$,

>>>>

\begin{example}\label{ex:SWT}
We build a small \SWT over the alphabets $\Sigma$ and $\Deltai$
(Ex.~\ref{ex:running} and \ref{distance-time}),
%weighted transducer model
with two states $q_0$ and $q_1$, that
calculates the temporal distance between an input performance in~$\Sigma^*$
and the subsequence
of $\Deltai$ events in a score.
%the performance $I$ and
%the score $O$ of Example~\ref{ex:running}.
%
Given  a performed event $\mu$
and the corresponding notated event $\nu$
(\eg MIDI pitch 69 and note $\mbox{\footnotesize A4}$),
the weight  computed by the \SWT is the time distance between both,
as modeled by transitions $\wei_{11}$ below.
%
The continuation symbol $'-'$
(met for instance in \emph{ties}
$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
or \emph{dots} \musQuarterDotted{})
%that represents continuation
is  skipped with no cost (transitions $\wei_{01}$).
%Likewise, the markups are ignored.
\[
\begin{array}{lclclcll}
\wei_{11}(q_0, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) & \quad & %|\tau' - \tau| & \quad &
\wei_{11}(q_1, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) &
     \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
\wei_{01}(q_0, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one\\
\wei_{10}(q_0, \mu\ts{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also want to take performing errors into account,
since a performer could, for example, play an unwritten extra note.
%\lydia{reformulated this sentence}
%
%Reading such a note is modelled by
The transition $\wei_{10}$,
with a fixed weight value $\alpha \in \Semiring$,
switches from state $q_0$ (normal) to $q_1$ (error)
when reading an extra note $\mu$.
%\philippe{Comprends pas cette phrase}
%\florent{$\to$ reformulation}
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and
$\init(q_1) = \final(q_1) = \zero$.
%\florent{dernière phrase utile?}
%This \SWT is capable of evaluating the differences between a score and a performance,
%all the while ensuring that performance errors are plausible.
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

@@

\noindent
 \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined  as the transducers of Definition~\ref{def:SWT},
by simply omitting the output symbols.
%
%In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$,
, respectively $\mathsf{out} : Q \to \Semiring$,  %\Phi_\Sigma,$
are functions defining the weight for entering
(respectively leaving) computation in a state,
and ${\wei_1}$ is a transition function
from $Q \times Q$ into~$\Phi_{\Sigma}$.
\end{definition}
%
\noindent
%\florent{mv formal def. weight for \SWA to app.?}
As above in the case of \SWT,
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$,
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$.
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$,
The computation of $A$ on words $s \in \Sigma^*$
is based on an intermediate function $\weight_A$,
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$
%
\begin{align}
\weight_A(q, \varepsilon, q') & = \one\ \mathrm{if}\  q = q'\ 
          \mathrm{and}\  \zero\  \mathrm{otherwise}\\
%\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q}
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$
\begin{equation}
A(s)  =
\displaystyle\bigoplus_{q, q' \in Q} \init(q)
\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q')
\label{eq:SWA-value}
\end{equation}

@@

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}\label{sec:SWVPA}
The model presented now generalizes symbolic \VPA
(\SVPA~\cite{dAntonyAlur14SVPDA},
 themselves generalizing \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
from Boolean semirings to arbitrary semiring domains.
It associates to every nested word over an infinite alphabet a weight value
in a semiring.
Nested words can describe structures of labeled trees.
In the context of parsing, they will be useful to
represent AST % parse trees.
(see Section~\ref{sec:parsing} and Appendix~\ref{app:trees}). % parse trees. 
%\florent{see \S 5 and App.A}

@@

\noindent
%The intuition is the following for the above transitions.
$\weiei$, $\weiec$, and $\weier$ describe the cases where the stack is empty.
%\lydia{moved this to the beginning}
%
\noindent
$\weii$ and $\weiei$ both read an input internal symbol $a$ and change state from $q$ to $q'$,
without changing the stack.
Moreover, $\weii$ reads a pair made of
${\call{c}} \in \Deltac$ and $p \in P$ on the top of the stack
($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$).
%
\noindent
$\weic$ and $\weiec$ read the input call symbol $\call{c}'$,
push it to the stack along with $p'$, and change state from $q$ to to $q'$.
Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack
($c$ is compared to $c'$).
%and $\weiec$ applies iff the stack is empty.
%
\noindent
$\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
Moreover, $\weir$ reads and pops from stack a pair made of $\call{c}$ and $p$,
($\call{c}$ is compared to $\return{r}$).
%and $\weier$ applies iff the stack is empty.
%In this case, the weight function $\phi_\mathsf{r}$
%computes a value of matching between the call and return symbols $c$ and $r$.
%This value might be set to $\zero$ in order to express that the symbols do not match.

Formally, the transitions of the automaton~$A$ are defined
with an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
%
A configuration $q[\gamma]$
is composed of a state $q \in Q$
and a stack content $\gamma \in \Gamma^*$,
where $\Gamma = \Deltac \times P$.
Hence, $\weight_A$ is a function from
$[Q \times \Gamma^*] \times \Delta^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
The empty stack is denoted by $\bot$, and the topmost
symbol is the last pushed pair.
%
The recursive definition of $\weight_A$
enumerates each of the six possible cases:
reading $a \in \Deltai$,
%\lydia{intro to func}
or $\call{c} \in \Deltac$, or $\return{r} \in \Deltar$,
for each possible state of the stack (empty or not).
%to add to $u \in {\Delta}^*$.
%\lydia{introduced the 6 cases}
\florent{shorter notation $cp$ for $\< c, p>$ ?}

@@

\noindent
Every $\SWA$ $A = \< Q, \init, {\wei_1}, \final >$,
over $\Sigma$, $\Semiring$ and $\bar\Phi$
is a special case of $\SWVPA$
$\< Q, \emptyset, \init, \bar\wei, \final >$
over $\Delta$, $\Semiring$ and $\bar\Phi$
with $\Deltai = \Sigma$ and $\Deltac = \Deltar = \emptyset$,
and computing with an always empty stack:
$\weiei = \wei_1$ and all the other functions
of~$\bar\wei$ are the constant~$\zero$.

@@

%Let us assume a third component in the symbols of $\Deltac$
%\philippe{Je trouve l'exemple un peu confus. Je ne sais pas
%quel est "l'associate time interval" d'un symbole. Je ne sais
%pas à quoi fait référence "the following call transition". Je
%n'ose pas trop toucher mais je crois qu'il faut revoir.}
%which is the duration of the associated time interval.
%It is used to check the timestamps in the symbols.
%
For instance, the following call transition starts in state
$q_{{i}/{c}}$, meaning that the current interval has number~$i < c$
amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the interval at the stack's top).
It reads a new time-division symbol~$\ccall{d}$:
\(
\weic\bigl(q_{{i}/{c}},
           \ccall{c}\ts{[\tau, \tau+\ell]}, q,
           \ccall{d}\ts{\iota}, q_{{i+1}/{c}},
           q_{{0}/{d}} \bigr) = \alpha_d
\).
The time interval $\iota$ attached to the~$\ccall{d}$ read 
must have a starting time $\tau + \frac{i\ell}{c}$,
where~$\tau$ and~$\ell$ are respectively
the starting time and duration of the interval attached to
the previous time-division symbol read $\ccall{c}$ (found on the stack).
%plus the total duration of intervals read so far, $i \ell$, and
%the new current interval duration is~$\frac{\ell}{d}$.
%
And it must have a duration $\frac{\ell}{d}$.

@@

%\subsection{Properties}
\medskip\noindent
Similarly to \VPA~\cite{AlurMadhusudan09nested}
and \SVPA~\cite{dAntonyAlur14SVPDA},
the class of \SWVPA is closed under the binary operators of the underlying semiring.
%
\begin{proposition}\label{prop:SWVPA-product}
Let $A_1$ and $A_2$ be two \SWVPA
over the same $\Delta$, commutative $\Semiring$ and effective $\bar\Phi$.
There exist two effectively constructible $\SWVPA$
$A_1 \oplus A_2$ and $A_1 \otimes A_2$,
such that for every $s \in \Delta^*$,
$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and
$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
\end{proposition}
%
\begin{proof}
We do a classical product construction, see Appendix~\ref{sec:product} for details.
%similar to the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
\end{proof}
\noindent
We present now a procedure for searching, for a~$\SWVPA$ $A$,
a word %$w \in \Delta^*$
of minimal weight for~$A$.
%as stated in the following proposition. %\wrt~$\leq_\oplus$,
%It requires the following property of label theories.
%
\begin{proposition}\label{th:best-search}
For a \SWVPA $A$
over $\Delta$,
$\Semiring$ commutative, bounded, total and complete, %semiring
and $\bar\Phi$ effective, %and $k$-convex, % label theory,
one can construct in PTIME a word $t \in \Delta^*$
such that $A(t)$ is minimal \wrt the natural ordering $\leq_\oplus$ for $\Semiring$.
\end{proposition}
%\florent{total?}
%
Let $A = \< Q, P, \init, \bar\wei, \final >$.
%
We propose a Dijkstra-like algorithm computing,
for every $q, q' \in Q$,
the minimum, \wrt $\leq_\oplus$, of the function
$\beta_{q, q'} : t \mapsto \weight_A(\config{q}{\bot}, t, \config{q'}{\bot})$.
Let us denote by $b_\bot(q, q')$ this minimum.
By definition of $\leq_\oplus$, and since $\Semiring$ is total,
it holds that:
%
\begin{equation}\label{eq:bbot}
  b_\bot(q, q') = \bigoplus_{t\in \Delta^*}
  \textstyle
  \weight_A\bigl(\config{q}{\bot}, t, \config{q'}{\bot}\bigr).
\end{equation}
The infinite sum in~\eqref{eq:bbot} is well defined since $\Semiring$ is complete.
%
Following~\eqref{eq:SWVPA-value}, and the associativity, commutativity
and distributivity for~$\otimes$ and~$\oplus$, the minimum of~$A(t)$ is:
\begin{equation}\label{eq:min}
{\displaystyle\bigoplus_{t\in \Delta^*}} A(t)
=
{\displaystyle\bigoplus_{t\in \Delta^*}}
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes}
\beta_{q, q'}(t)
\mathop{\otimes} \mathsf{out}(q')
=
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes}
b_\bot(q, q')
\mathop{\otimes} \mathsf{out}(q')
\end{equation}

\noindent
In order to compute the above function $b_\bot : Q \times Q \to \Semiring$,
we shall consider an auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
%
Intuitively, $b_\top(q, p, q')$ is a function of $\Phic$,
mapping every $c \in \Deltac$ to
the minimum weight of a computation of~$A$
starting in state $q$ with a non-empty stack
$\gamma' = \<c, p> \, \gamma \in \Gamma^+$,
and ending in state $q'$ with the same stack $\gamma'$,
such that moreover,
the computation does not pop the pair $\<c, p>$ at the top of $\gamma'$
(\ie $\gamma'$ is left untouched during the computation).
However, the computation can read $\<c, p>$ at the top of $\gamma'$,
and can also push another pair $\< c', p'> \in \Gamma$ on $\gamma'$,
following the third case 
in the definition \eqref{eq:SWVPA-weight} of $\weight_A$ (call symbol).
The pair $\< c', p'>$ can be popped later, during the computation from $q$ to $q'$,
following the fifth case of \eqref{eq:SWVPA-weight} (return symbol).
%However, it cannot apply one of the two last cases (return symbol and empty stack)
%when the current stack is $\gamma$.
%pop symbols in $\gamma$.
%
% Note that having a stack reduced to such a symbol makes impossible the application of the
% two last cases in the definition of $\weight_A$ (return symbol and empty stack).
% However, it is possible to apply the two first cases
% (internal symbol or call symbol, with a push on the top of $\top$).
%
Formally, in order to define $b_\top$, we consider
a fresh stack symbol $\top \notin \Gamma$,   %which does not belong to $\Gamma$,
representing the above untouched stack, and let:
%
\begin{equation}\label{eq:btop}
  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Delta^*}
  \textstyle
  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
\quad
\mathrm{~for~all~} c \in \Deltac
\end{equation}
%
By definition of $\weight_A$ in \eqref{eq:SWVPA-weight},
using the symbol $\top$ for the part of the stack below $\<c, p>$
(\ie the substack $\gamma$ in the above $\gamma' = \<c, p> \, \gamma$)
ensures that this part is not touched during the computation.
% there cannot be a pop of $\<c, p>$ followed by a push of $\<c, p>$
% because it is not possible to push over $\top$ (push read the top of stack).
%
This ensures in particular that the subword
read during the computation is well parenthesized
(every symbol in $\Deltac$ has a matching symbol in $\Deltar$).

@@

Let us now apply the models and results of the previous sections %in order to define
to the problem of parsing over an infinite alphabet. %appropriate
%
%\subsection{Definition}
%
Let~$\Sigma$
and~$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$
be countable input and output alphabets,
let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a
commutative, bounded, total and complete %\florent{total?} 
semiring and let $\bar\Phi$ be an effective label theory over $\Semiring$,
containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Deltai}$, as well as
$\Phii$, $\Phic$, $\Phir$, $\Phicr$
(following the notations of Section~\ref{sec:SWVPA-def}).
%
\noindent
We assume the following input to be given:
\begin{description}
\item[--] a \SWT $T$ over $\Sigma$, $\Deltai$, $\Semiring$, and $\bar\Phi$,
defining a measure %between words
$T: \Sigma^* \times \Deltai^* \to \Semiring$,
